---
title: "Accessing the Allele Frequency Net Database using *AFNDquery*"
author:
- name: Katharina Imkeller
  affiliation: European Molecular Biology Laboratory, Heidelberg
  email: imkeller@embl.de
output:
  html_document
---

```{r}
library(AFNDquery)
```

## Introduction

The Allele Frequency Net Database (http://www.allelefrequencies.net/) is a repository for immune gene frequencies in different populations worldwide. This package allows the automated R access to the HLA related data within the Allele Frequency Net Database. The query principle of this package currently relies on parsing information from the html-formated pages of the online resource.

Citing the Allele Frequency Net Database:\\Allele frequency net 2015 update: new features for HLA epitopes, KIR and disease and HLA adverse drug reaction associations. Gonzalez-Galarza FF, Takeshita LY, Santos EJ, Kempson F, Maia MH, Silva AL, Silva AL, Ghattaoraya GS, Alfirevic A, Jones AR and Middleton D Nucleic Acid Research 2015, 28, D784-8.

The HLA frequency datasets in the Allele Frequency Net Database are classified according to the following standards:

| Criteria        | Gold standard           | Silver standard  | Bronze standard |
| ----------- |:-----------------:|:-----------------:|:----------:|
| Allele frequency      | Sum to 1 (± 0.015; or ± 0.05) | Sum to 1 (± 0.015; or ± 0.05) | other |
| Sample size     | >= 50 individuals      |   any | any |
| Resolution | four or more digits      | two or more digits | other |

## Scope of the AFNDquery package

The pAFNDquery package currently supports the following type of queries:

1. Query the **HLA allele frequencies** for a given HLA locus, population, sample size (and other available filters). This module returns information contained in the *HLA* section of the Allele Frequency Net Database (allele frequency, population name and id, sample size). The queries can basically be tha same as the ones entered into the online user interface at http://www.allelefrequencies.net/hla6006a.asp

2. Query the **population metainformation** for a given population. This module returns information contained in the *Population* section of the Allele Frequency Net Database (ethnic origin, geographic location, sampling size and date, ...). For the HLA allele dataset of a population in Smolensk, Russia, this page can be found here: http://www.allelefrequencies.net/pop6001c.asp?pop_id=0002645 

## Querying allele frequencies

**Example 1:** Querying the freqeuncy of allele A*02:01 in all populations with more than 10,000 individuals and fulfilling the gold standard.

```{r}
sel1 <- query_allele_frequencies(hla_selection = "A*02:01", 
                                hla_sample_size_pattern = "bigger_than", 
                                hla_sample_size = 10000, 
                                standard="g")

DT::datatable(sel1)
```

**Example 2:** Querying the frequencies of alleles within the HLA-B locus for population "Peru Lamas City Lama" with population_id 1986. If you are not sure which population_id you should use, we recommend to search your population of interest in the *Population* section of the Allele Frequency Net Database and extract the population id from the url (last 4 digits).

```{r}
sel2 <- query_allele_frequencies(hla_locus = "B", hla_population = 1986)

DT::datatable(sel2)
```

## Querying population metainformation

**Example 3:** Query the metainformation concerning population "Peru Lamas City Lama" (population_id 1986). The webpage concerning the queried information for population Peru Lamas City Lama (1986) can be found here: http://www.allelefrequencies.net/pop6001c.asp?pop_id=1986

```{r}
sel3 <- query_population_detail(1986)

DT::datatable(sel3, options = list(scrollX = TRUE))
```

**Example 4:** Query the metainformation concerning the populations that were listed in the table returned by Example 1

```{r}
sel4 <- query_population_detail(as.numeric(sel1$population_id))

DT::datatable(sel4, options = list(scrollX = TRUE))
```


